---
title: "Group Task"
author: "Phong Duong"
format:
  html:
    theme: [style.scss]
    toc: true
    toc-location: right
    toc-title: Ãndice
editor: visual
---

# Required packages

> Insert in the lower chunk the packages you will need

```{r}
rm(list = ls()) 
library(tidyverse)
library(dbplyr)
library(RSQLite)
library(DBI)
```

# Data

```{r}
db <- DBI::dbConnect(SQLite(), "database.sqlite", extended_types = T)
dbListTables(db)
```

# Data Processing

```{r}
tbl(db, "election_data") %>% 
  head(n = 50)
```

# Questions

1.  How is the vote of national parties (PSOE, PP, VOX, CS, MP, UP - IU) distributed against regional or nationalist parties?

```{r}

```

2.  Which party was the winner in the municipalities with more than 100,000 habitants (census) in each of the elections?

```{r}

```

3.  Which party was the second when the first was the PSOE? And when the first was the PP?

```{r}

```

4.  Who benefits from low turnout?

```{r}

```

5.  How to analyze the relationship between census and vote? Is it true that certain parties win in rural areas?

```{r}

```

6.  How to calibrate the error of the polls (remember that the polls are voting intentions at national level)?

```{r}

```

7.  In which election were the polls most wrong?

```{r}

```

8.  How were the polls wrong in national parties (PSOE, PP, VOX, CS, MP, UP - IU)?

```{r}

```

9.  Which polling houses got it right the most and which ones deviated the most from the results?

```{r}

```
