---
title: "Group Task"
author: "Valeria Contreras-Phong Duong-Jieun Park-Ana Pérez-Sofia Villamil"
format:
  revealjs:
    theme: moon
    incremental: true
    width: 1920
    height: 1080
editor: visual
---

## Required packages {visibility="hidden"}

Insert in the lower chunk the packages you will need

```{r}
rm(list = ls()) 
library(tidyverse)
library(dbplyr)
library(RSQLite)
library(DBI)
library(lubridate)
library(scales)
library(glue)
library(ggplot2)

```

## Required packages

-   tidyverse
-   dbplyr
-   RSQLite
-   DBI
-   lubridate
-   scales

# About the data

## Original data {visibility="hidden"}

```{r}
if(!file.exists("database.sqlite")) {
  #Original data files
  election_data <- read_csv(file = "./data/datos_elecciones_brutos.csv")
  cod_mun <- read_csv(file = "./data/cod_mun.csv")
  surveys <- read_csv(file = "./data/historical_surveys.csv")
  abbrev <- read_csv(file = "./data/siglas.csv")
  
  #Pivoted election_data table name: election_data
  election_df <- election_data %>% 
    pivot_longer(cols = !tipo_eleccion:votos_candidaturas, names_to = 'party', values_to = 'vote_count') %>% 
    select(-tipo_eleccion, -vuelta, -codigo_distrito_electoral) %>% #column with 0 variance
    mutate(vote_count = replace_na(vote_count, 0),
           cod_mun = glue('{codigo_ccaa}-{codigo_provincia}-{codigo_municipio}'), .after = codigo_municipio) 
  
  #Municipal population table
  municipal_population <- election_data %>% 
    mutate(cod_mun = glue('{codigo_ccaa}-{codigo_provincia}-{codigo_municipio}'), .after = codigo_municipio) %>% 
    select(anno, mes, cod_mun, censo) %>% 
    summarise(sum(censo), .by = c(cod_mun, anno, mes)) %>% 
    rename('censo_total' = 'sum(censo)')
  
  #Pivoted surveys table name: surveys
  surveys_df <- surveys %>% 
    pivot_longer(cols = !type_survey:turnout, names_to = 'siglas', values_to = 'estimated_voting_intention') %>% 
    mutate(date_elec = as.character(date_elec),
           field_date_from = as.character(field_date_from),
           field_date_to = as.character(field_date_to)) 
  
  #Migrate data to sqlite database
  db_file <- "database.sqlite"
  src_sqlite(db_file, create = TRUE) #Create database
  db <- DBI::dbConnect(SQLite(), "database.sqlite", extended_types = T) #Connect to sqlite database
  
  dbWriteTable(db, "election_data", election_df, field.types = c(anno = "Int"))
  dbWriteTable(db, "municipal_population", municipal_population)
  dbWriteTable(db, "cod_mun", cod_mun)
  dbWriteTable(db, "surveys", surveys_df, field.types = c(date_elec = "Date", field_date_from ="Date", field_date_to = "Date"))
  dbWriteTable(db, "abbrev", abbrev)
  
  #Filtered surveys table for ease of querying
  dbSendQuery(db, "CREATE TABLE surveys_new AS
  SELECT *
  FROM surveys
  WHERE date_elec >= '2018-01-01'
   AND exit_poll = 0
   AND size >= 750
   AND JULIANDAY(field_date_to) - JULIANDAY(field_date_from) > 1;
  ")
  
  #Add missing parties in abbrev table 
  dbSendQuery(db, "INSERT INTO abbrev(denominacion)
  SELECT DISTINCT party
  FROM election_data 
  WHERE party NOT IN (SELECT denominacion FROM abbrev);")
  
  #Processed party names and abbreviation
  dbSendQuery(db, "CREATE TABLE abbrev_2 AS
  SELECT DISTINCT denominacion, 
      CASE 
          WHEN siglas LIKE '%PSOE%' OR siglas LIKE 'P.S.O.E.' OR denominacion LIKE '%PSOE%' THEN 'PSOE'
          WHEN siglas LIKE '%P.P.%' OR denominacion LIKE '%PARTIDO POPULAR%' THEN 'PP'
          WHEN denominacion LIKE '%PARTIDO DE LA CIUDADAN_A%'  THEN 'CS'
          WHEN denominacion LIKE '%VASCO%' THEN 'EAJ'
          WHEN denominacion LIKE '%NACIONALISTA GALEGO%' THEN 'BNG'
          WHEN siglas LIKE '%COMPRO%' THEN 'COM'
          WHEN denominacion LIKE '%UNIÓ' OR siglas = 'CIU' THEN 'CIU'
          WHEN denominacion LIKE '%ESQUE    RRA REPUBLICANA%' AND denominacion NOT LIKE '%VALENCI%' THEN 'ERC'
          WHEN denominacion LIKE '%BILDU%'  OR denominacion LIKE '%Aralar%' OR denominacion LIKE '%Eusko%' THEN 'EH-BILDU'
          WHEN denominacion LIKE '%los verdes%' OR denominacion LIKE '%podem%' OR denominacion LIKE '%ezker batua%' OR denominacion LIKE '%IZQUIERDA UNIDA%' THEN 'UP-IU'
          WHEN denominacion LIKE '%MÁS PAÍS%' THEN'MP'
          WHEN denominacion LIKE '%VOX%' THEN 'VOX'
          ELSE 'OTHER'
      END AS siglas_2
  FROM abbrev;")
  
  #New election data filtered and processed date
  dbSendQuery(db, "CREATE TABLE election_data_2 AS
  SELECT 
      (anno || '-' || mes || '-' || '01') AS elec_date,
      cod_mun,numero_mesas, participacion_1, participacion_2,
      votos_blancos, votos_nulos, votos_candidaturas, party, vote_count
  FROM election_data;")
  
  rm(list=ls())
}
```

## Data

```{r}
#| echo: true
db <- DBI::dbConnect(SQLite(), "database.sqlite", extended_types = T)
```

Due to the large size of the pivoted **`election_data`** dataframe (22 million rows), handling it became computationally intensive. To optimize efficiency, we migrated the data to a SQL database, enabling selective retrieval of necessary tables and reducing the load on the R environment.

# The Analysis

## Distribution of national parties vs regional

1.How is the vote of national parties (PSOE, PP, VOX, CS, MP, UP - IU) distributed against regional or nationalist parties?

```{r}
#Create a table in sql db for faster data retrieval
if(!dbListTables(db) %>% str_detect("vote_sum") %>% any()) {
  q1_df <- tbl(db, "election_data_2") %>% 
  select(cod_mun, elec_date, party, vote_count) %>% 
  left_join(tbl(db, "abbrev_2"), by = join_by("party" == "denominacion")) %>% 
  mutate(nat_reg = ifelse(siglas_2 %in% c('PSOE', 'PP','VOX', 'CS', 'MP', 'UP-IU'),'national', 'regional'))
  
  vote_sum <- q1_df %>% 
    group_by(nat_reg, elec_date) %>% 
    summarise(vote_total = sum(vote_count)) 
  vote_sum %>% head()
  compute(vote_sum, temporary = FALSE, name='vote_sum')
}
```

::: columns
::: {.column width="50%" style="align: center"}
```{r}
#| fig-height: 7
#| fig-width : 13
plot1_data <- tbl(db, "vote_sum") 
ggplot(plot1_data, aes(x = elec_date, y = vote_total/10^6, fill = nat_reg)) +
  geom_col(position = position_dodge2()) +
  labs( title = "Vote Distribution: National vs. Regional/ Nationalist Parties",
  subtitle = "Analysis of votes for national parties (PSOE, PP, VOX, CS, MP, UP-IU) compared to regional or nationalist parties",
  x = "Election Date",
  y = "Total Votes (in millions)",
  fill = NULL
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )



```
:::

::: {.column width="30%" style="align: center"}
### Table

```{r}
tbl(db, "vote_sum") %>% collect() %>% tibble()
```
:::
:::

::: {style="text-align: center"}
National parities largely outweigh regional parties in votes. However, notable surges in votes for regional parties during certain elections, such as in 2011, suggest there were impactful events occurring between elections.
:::

## Dominant Political Parties in Municipalities

2.Which party was the winner in the municipalities with more than 100,000 habitants (census) in each of the elections?

```{r}
#Create another table for ease of computing
if(!dbListTables(db) %>% str_detect("vote_full") %>% any()) {
    vote_full <- tbl(db, "election_data_2") %>% 
      group_by(elec_date, cod_mun) %>% 
      mutate(valid_votes = sum(vote_count) + votos_blancos,
             percent = vote_count/ valid_votes * 100) %>% 
    select(elec_date, cod_mun, party, vote_count, valid_votes, percent)
    compute(vote_full, temporary = FALSE, name = "vote_full")
}
```

```{r}
mun_pop_2 <- tbl(db, "municipal_population") %>% 
  mutate(date = paste(as.integer(anno), mes, "01", sep = "-" )) %>% 
  select(-anno, -mes) 

winner_100k <- tbl(db, "vote_full") %>% 
  left_join(mun_pop_2, by = join_by("cod_mun", "elec_date" == "date")) %>% 
  filter(censo_total > 100000) %>% 
  group_by(elec_date, cod_mun) %>% 
  filter(percent == max(percent)) %>% 
  ungroup() %>% 
  collect()
```

```{r}
#Helper function to look for winner of a certain year
winner <- function(elec) {
 return(
   winner_100k %>% 
     filter(elec_date == elec) %>% 
     group_by(party) %>% 
     summarise(wins = n()) %>% 
     slice_max(wins, n = 1)
 ) 
}
```

```{r}
#| output: false

#2008
winner('2008-03-01')

#2011
winner('2011-11-01')

#2015
winner('2015-12-01')

#2016
winner('2016-06-01')

#2019-1
winner('2019-04-01')

#2019-2
winner('2019-11-01')
```

```{r}
#Summary data frame to make the plot
winner_100k_summary <- winner_100k %>% 
  group_by(elec_date, party) %>% 
  summarise(wins = n()) %>% 
  filter(wins == max(wins)) %>% 
  left_join(tbl(db, "abbrev_2") %>% collect(), by = join_by("party" == "denominacion")) 
```

::: {style="text-align: center"}
```{r}
party_colors <- c("PP" = "#0157a1", "PSOE" = "#f20400")

ggplot(winner_100k_summary) +
  geom_col(aes(x = elec_date, y = wins, fill = siglas_2)) +
  geom_text(aes(x = elec_date, y = wins, label = siglas_2), nudge_y = 1) +
  scale_fill_manual(values = party_colors) +  
  labs(
    title = "Election Winners in Cities Over 100k Inhabitants",
    subtitle = "Comparative wins of PP and PSOE across multiple elections",
    x = "Election Date",
    y = "Number of Wins",
    fill = "Party"
  ) + 
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12)
  )

```
:::

::: {style="text-align: center"}
In larger municipalities, PP and PSOE predominantly emerge as the winning parties
:::

## Primary Runner-Up Parties to PSOE and PP in Elections

3.Which party was the second when the first was the PSOE? And when the first was the PP?

```{r}

first_second <- tbl(db, "vote_full") %>% 
  left_join(mun_pop_2, by = join_by("cod_mun", "elec_date" == "date")) %>% 
  filter(censo_total > 100000) %>% 
  relocate(censo_total, .after = cod_mun) %>% 
  left_join(tbl(db, "abbrev_2"), by = join_by("party" == "denominacion")) %>% 
  relocate(siglas_2, .after = party) %>% 
  group_by(elec_date, cod_mun) %>% 
  arrange(elec_date, cod_mun) %>% 
  slice_max(vote_count, n = 2) %>% 
  collect()
```

```{r}
#| output: false
#First is PSOE

#Count how many times got second place
first_second %>% 
  filter(siglas_2[1] == "PSOE") %>% 
  filter(siglas_2 != "PSOE") %>% ungroup() %>% 
  select(elec_date, party, siglas_2, percent) %>% 
  group_by(elec_date, siglas_2) %>% 
  summarise(count = n()) %>% 
  arrange(elec_date, desc(count))

#Sum votes per election of second place party
first_psoe <- first_second %>% 
  summarise(first = siglas_2[1],
            second = siglas_2[2],
            
             votes_first = sum(vote_count[siglas_2 == first]),
             votes_second = sum(vote_count[siglas_2 == second]),
            .groups = 'drop' ) %>% 
   
  filter(first== "PSOE") %>% 
  ungroup() %>% 
   
  group_by(elec_date, second, first) %>% 
  summarise(count_second = sum(votes_second), 
            count_first = sum(votes_first), 
            .groups = 'drop') %>% 
 group_by(elec_date) %>% 
 filter(count_second == max(count_second))
```

```{r}
ggplot(first_psoe, aes(y = reorder(elec_date, desc(elec_date)), 
                     x = count_second, fill = second)) +
  geom_bar(stat = "identity", position = "identity", width = 0.7) +
  scale_fill_manual(values = c("PP" = "#0157a1", "UP-IU" = "#663278")) +

  labs(x="", y="", title= "Second party when the winner is PSOE", subtitle = "total nº of votes per election in municipalites with more than 100.000 inhabitants") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle = element_text(face = "italic", size = 12, hjust = 0),
    legend.position = "bottom",
    legend.title=element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title = element_blank(),
    panel.grid.major.y = element_line(color = "gray", linetype = "dashed", size = 0.3)
  )  
  
```

```{r}
#| output: false

#First is PP
#Count how many times got second place
first_second %>% 
  filter(siglas_2[1] == "PP") %>% 
  filter(siglas_2 != "PP") %>% ungroup() %>% 
  select(elec_date, party, siglas_2, percent) %>% 
  group_by(elec_date, siglas_2) %>% 
  summarise(count = n()) %>% 
  arrange(elec_date, desc(count))

#Sum votes per election of second place party
first_pp <- first_second %>% 
  summarise(first = siglas_2[1],
            second = siglas_2[2],
            
             votes_first = sum(vote_count[siglas_2 == first]),
             votes_second = sum(vote_count[siglas_2 == second]),
            .groups = 'drop' ) %>% 
   
  filter(first== "PP") %>% 
  ungroup() %>% 
   
  group_by(elec_date, second, first) %>% 
  summarise(count_second = sum(votes_second), 
            count_first = sum(votes_first), 
            .groups = 'drop') %>% 
 group_by(elec_date) %>% 
 filter(count_second == max(count_second))
```

```{r}
ggplot(first_pp, aes(y = reorder(elec_date, desc(elec_date)), 
                     x = count_second, fill = second)) +
  
  geom_bar(stat = "identity", position = "identity", width = 0.7) +
  scale_fill_manual(values = c("PSOE" = "#f20400", "UP-IU" = "#663278")) +
  theme_minimal() +
  labs(x="", y="", title= "Second party when the winner is PP", subtitle = "total nº of votes per election in municipalites with more than 100.000 inhabitants") +
   theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle = element_text(face = "italic", size = 12, hjust = 0),
    legend.position = "bottom",
    legend.title=element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title = element_blank(),
    panel.grid.major.y = element_line(color = "gray", linetype = "dashed", size = 0.3)
  ) 
```

## Impact of Voter Turnout on Election Outcomes

4.Who benefits from low turnout?

```{r}
tbl(db,"election_data_2")

#Long run time, consider optimizing 
lowturnout <- tbl(db,"election_data_2") %>% 
  left_join(tbl(db, "abbrev_2"),by = join_by("party" == "denominacion")) %>% 
  left_join(mun_pop_2, by = join_by("cod_mun", "elec_date" == "date")) %>% 
  group_by(elec_date, cod_mun) %>% 
  mutate(participation_pcg = participacion_2/censo_total * 100) %>% 
  select(elec_date, cod_mun, party, siglas_2, participacion_2, vote_count,  participation_pcg, censo_total) %>% 
  filter(participation_pcg < 30) %>% 
  arrange(participation_pcg) %>% 
  filter(vote_count == max(vote_count)) 
```

```{r}
library(forcats)
ggplot(lowturnout, aes(x = fct_reorder(siglas_2,vote_count), 
                       y = vote_count, fill = siglas_2)) +
  
  geom_bar(stat = "identity", na.rm = TRUE) +
  scale_fill_manual(values = c("PP" ="#0157a1","PSOE" = "#f20400","EAJ" = "#00822a"))+
                    
  theme_minimal()+

coord_polar() +
  annotate( x = 11.5, y = 1100, label = "1000",
            geom = "text", color = "gray12",size= 3, na.rm = TRUE) +
 
   annotate( x = 11.5,y = 2100, label = "2000", 
             geom = "text", color = "gray12", size=3, na.rm = TRUE) +
 
   annotate( x = 11.5, y =3100, label = "3000",
             geom = "text", color = "gray12", size=3, na.rm = TRUE) +
  
  annotate( x = 11.7,  y =4100, label = "4000", 
            geom = "text", color = "gray12", size=3, na.rm = TRUE ) +
  
 
   scale_y_continuous(
    limits = c(-1500, 3500),
    expand = c(0, 0),
    breaks = c(0, 1000, 2000, 3000))+
   
  theme(axis.text.y = element_blank(), 
        axis.text.x = element_text(color = "gray12", size = 8), 
        legend.position = "none", 
        legend.title = element_blank(), 
        plot.title = element_text(face="bold")) +
  
  labs( x="", y="", title = "\nTotal votes for each party in municipalities with less than 30% of turnout", 
            subtitle = paste(
      "\nPP and PSOE still the most benefited parties when low turnout,",
      "EAJ-PNV gains votes from low turnout.\n",
       sep = "\n"
))
```

### Plot

## Analyzing the Correlation Between Population Census and Voting Patterns

5.How to analyze the relationship between census and vote? Is it true that certain parties win in rural areas?

```{r}
#Vote data frame of interested parties and in rural area (population <= 5000)
#Criteria for rural based on DEMOGRAFÍA DE LA POBLACIÓN RURAL en 2020
rural_winners <- tbl(db, "vote_full") %>% 
  left_join(tbl(db, "abbrev_2"), by = join_by("party" == "denominacion")) %>% 
  filter(siglas_2 != "OTHER") %>% 
  left_join(mun_pop_2, by = join_by("cod_mun", "elec_date" == "date")) %>%
  filter(censo_total <= 5000) %>% 
  collect()

#Winning times in rural zones per party per elections 
rural_winners %>% 
  group_by(elec_date, cod_mun) %>% 
  filter(percent == max(percent)) %>% 
  group_by(elec_date, siglas_2) %>% 
  summarise(win_count = n()) %>% 
  arrange(elec_date,desc(win_count)) |> 
  filter(elec_date == "2008-03-01")
```

```{r}
#2008
rural_winners %>% 
  group_by(elec_date, cod_mun) %>% 
  filter(percent == max(percent) & elec_date == "2008-03-01") %>% 
  ggplot(aes(x = percent, fill = siglas_2)) +
  geom_histogram(binwidth = 1, show.legend = TRUE) +  
  labs(
    title = "Winning Times in Rural Zones per Party",
    subtitle = "Election Date: 2008-03-01",
    x = "Percentage of Votes",
    y = "Frequency",
    fill = "Party"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.text = element_text(size = 5),
    legend.key.size = unit(5, "mm")
  ) +
  scale_fill_manual(values = c(
    "PSOE" = "#f20400",
    "PP" = "#0157a1",
    "UP-IU" = "#663278",
    "CS" = "#e9550d",
    "EAJ" = "#00822a",
    "VOX" = "#76b829", 
    "MP" ="#00e1c5",
    "CIU" = "#f68b22",
    "EH-BILDU" = "#95c11f",
    "BNG" = "#76b4de",
    "COM" = "#8c3651"
  )) 
  
```

# Analyzing Polling Errors: Accuracy and Discrepancies in Election Forecasts

## Calibrating the error of the polls

6.How to calibrate the error of the polls (remember that the polls are voting intentions at national level - national parties: PSOE, PP, VOX, CS, MP, UP - IU)?

```{r}
#| output: false
# Obtaining the average estimated voting intention for each party
voting_intention <- tbl(db, "surveys") %>%
  filter(date_elec >= '2008-01-01') %>% 
  select(siglas, date_elec, field_date_from, field_date_to,estimated_voting_intention) %>%
  mutate(siglas_imp = case_when(siglas == "PSOE" ~ "PSOE",
                                siglas == "PP" ~ "PP",
                                siglas == "CS" ~ "CS",
                                siglas %in% c("PODEMOS", "UP", "IU") ~ "UP-IU",
                                siglas == "MP" ~ "MP",
                                siglas == "VOX" ~ "VOX",
                                TRUE ~ "OTHER"), .after = "siglas") %>% 
  group_by(date_elec, siglas_imp) %>%
  summarise(average_intention = mean(estimated_voting_intention)) %>% 
  collect()

# Calculate the national vote share for each party
vote_share_real <- tbl(db, "vote_full") %>% 
  left_join(tbl(db, "abbrev_2"), by = join_by("party" == "denominacion")) %>%  
  select(elec_date, cod_mun, party, siglas_2, vote_count, valid_votes) %>% 
  mutate(siglas_2 = ifelse(siglas_2 %in% c('PSOE', 'PP','VOX', 'CS', 'MP', 'UP-IU'),siglas_2, 'OTHER')) %>% 
  group_by(elec_date,cod_mun, siglas_2) %>% 
  summarise(total_votes = sum(vote_count), valid_votes) %>% 
  group_by(elec_date, siglas_2) %>% 
  summarise(national_share = sum(total_votes)/sum(valid_votes) *100) %>% 
  collect() 
```

```{r}
#| output: false
#Comparison(election 2019-11-01)
vote_share_real %>% 
  mutate(elec_date = ymd(elec_date)) %>% 
  filter(elec_date == '2019-11-01') %>% 
  left_join(voting_intention, by = join_by(closest("elec_date" <= "date_elec"),"siglas_2" == "siglas_imp")) %>% 
  mutate(poll_error = average_intention - national_share) %>% 
  ungroup() %>% 
  select(-elec_date) %>% relocate(date_elec, .before = siglas_2)
```

### Table

```{r}
table_data <- vote_share_real %>% 
  mutate(elec_date = ymd(elec_date)) %>% 
  filter(elec_date == '2019-11-01') %>% 
  left_join(voting_intention, by = join_by(closest("elec_date" <= "date_elec"),"siglas_2" == "siglas_imp")) %>% 
  mutate(poll_error = average_intention - national_share) %>% 
  ungroup() %>% 
  select(-elec_date) %>% relocate(date_elec, .before = siglas_2) %>%
  select(siglas_2, poll_error)
table_data

```

## Identifying the Election with the Greatest Polling Discrepancies

7.In which election were the polls most wrong?

```{r}
#| output: false
error_table <- vote_share_real %>% 
  mutate(elec_date = ymd(elec_date)) %>% 
  left_join(voting_intention, by = join_by(closest("elec_date" <= "date_elec"),"siglas_2" == "siglas_imp")) %>% 
  mutate(poll_error = average_intention - national_share) %>% 
  ungroup() %>% 
  select(-elec_date) %>% relocate(date_elec, .before = siglas_2) %>% 
  group_by(date_elec) %>% 
  summarise(total_error = sum(abs(poll_error), na.rm = TRUE)) 

error_table |>  slice_max(total_error, n =1) 

```

```{r}
ggplot(error_table, aes(x = as.character(date_elec), y = total_error)) +
  geom_bar(stat = "identity", fill = "#87CEEB") +
  labs(
    title = "Total Error by Election Date",
    x = "Election Date",
    y = "Total Error"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),  
    axis.title = element_text(face = "bold", size = 12),  
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 10),  
    axis.text.y = element_text(size = 10)) +
  scale_y_continuous(expand = expansion(mult = c(0, .1)))  

```

## Assessing Polling Inaccuracies for National Parties

8.How were the polls wrong in national parties (PSOE, PP, VOX, CS, MP, UP - IU)?

```{r}
error_by_party <- vote_share_real %>% 
  mutate(elec_date = ymd(elec_date)) %>% 
  left_join(voting_intention, by = join_by(closest("elec_date" <= "date_elec"),"siglas_2" == "siglas_imp")) %>% 
  mutate(poll_error = average_intention - national_share) %>% 
  ungroup() %>% 
  select(-elec_date) %>% relocate(date_elec, .before = siglas_2) %>% 
  filter(siglas_2 != "OTHER") %>% 
  group_by(siglas_2) %>%  
  summarise(total_error = sum(abs(poll_error), na.rm = TRUE))
```

```{r}
ggplot(error_by_party, aes(x = siglas_2, y = total_error)) +
  geom_bar(stat = "identity", fill = "#87CEEB" ) +  
  labs(
    title = "Total Error by Party",
    x = "Party",
    y = "Total Error"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),  
    axis.title = element_text(face = "bold", size = 12),  
    axis.text.x = element_text(angle = 45, hjust = 1),    
    axis.text.y = element_text(size = 10),                
    legend.position = "none"                              
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) 

```

### Table

```{r}
print(error_by_party)
```

## Identifying Top Performers and Those with Greatest Deviations

9.Which polling houses got it right the most and which ones deviated the most from the results?

```{r}
#| output: false
pollster_estimate <- tbl(db, "surveys") %>%
  filter(date_elec >= '2008-01-01') %>% 
  select(siglas, pollster, date_elec, field_date_from, field_date_to, 
         estimated_voting_intention) %>%
  mutate(siglas_imp = case_when(siglas == "PSOE" ~ "PSOE",
                                siglas == "PP" ~ "PP",
                                siglas == "CS" ~ "CS",
                                siglas %in% c("PODEMOS", "UP", "IU") ~ "UP-IU",
                                siglas == "MP" ~ "MP",
                                siglas == "VOX" ~ "VOX",
                                TRUE ~ "OTHER"), .after = "siglas") %>% 
  group_by(date_elec, pollster, siglas_imp) %>%
  summarise(average_intention = mean(estimated_voting_intention)) %>% 
  collect() %>% 
  mutate(winner_estimated = siglas_imp[which.max(average_intention)])


pollster_result <- pollster_estimate %>% 
  left_join(vote_share_real %>%
              mutate(elec_date = ymd(elec_date),
                     winner = siglas_2[which.max(national_share)]), 
            by = join_by(closest("date_elec" >= "elec_date"),
                         "siglas_imp" == "siglas_2"), 
            keep = TRUE) %>% 
  mutate(correct_guess = ifelse(winner_estimated == winner, 1,0),
         error = average_intention - national_share) %>% 
  select(-elec_date)
```

```{r}
#| output: false
#How many times pollsters correctly estimated winning party
correct_estimates <- pollster_result %>% 
  select(date_elec, pollster, correct_guess) %>% 
  distinct() %>% 
  group_by(pollster) %>% 
  summarise(correct_guesses = sum(correct_guess))

#top pollsters
correct_estimates %>% slice_max(correct_guesses)

#worst pollsters
correct_estimates %>% slice_min(correct_guesses)
```

```{r}
#| output: false
summarise_pollster_error <- pollster_result %>% 
  select(date_elec, pollster, error) %>% 
  group_by(pollster) %>% 
  summarise(avg_error = mean(abs(error), na.rm = TRUE))
  
#Best pollster
summarise_pollster_error %>% slice_min(avg_error)

#Worst pollster
summarise_pollster_error %>% slice_max(avg_error)
```

```{r}
ggplot(summarise_pollster_error, aes(x = reorder(pollster, avg_error), y = avg_error)) +
  geom_point(color = "#87CEEB") +  
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.major = element_line(color = "grey80", size = 0.5),  
    panel.grid.minor = element_blank(), 
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(face = "bold", size = 12),
    axis.title.x = element_text(face = "bold", size = 10),
    axis.title.y = element_text(face = "bold", size = 10)
  ) +
  labs(
    title = "Average Polling Error by Pollster",
    subtitle = "Comparison of polling errors across different organizations",
    x = "Pollster",
    y = "Average Error"
  ) +
  scale_y_continuous(labels = scales::comma)
```

------------------------------------------------------------------------

::: {style="text-align: center; font-size: xx-large; font-weight: bold;"}
THE END
:::

::: {style="text-align: center; font-size: larger;"}
<strong>We are open for any questions you guys have.</strong>
:::
